The role of cognateness in non-native spoken word recognition

### 2021-12-04 Rmd/manuscript.Rmd
knit to Word

bibliography:
  - r-references.bib
  - references.bib
floatsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: no
draft: no
csl: "apa7.csl"
documentclass: apa7
classoption: man
appendix: "manuscript_appendix.Rmd"
output:
  word_document:
    reference_docx: "manuscript_template.docx"
---

@@

options(knitr.kable.NA = '-', knitr.duplicate.label = "allow")

@@

@@

# Appendix

```{r accuracy, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Proportion of correct translation by item. Words presented to participants in the English\\-Spanish or in the English\\-Catalan and Spanish\\-Catalan are listed in the Y\\-axis, ordered from higher to lower average translation accuracy, depicted in the X\\-axis. Dots and whiskers represent the average accuracy and 95\\% confidence interval of each word. Accuracy is plotted separately for each group.", fig.height=10, fig.width=9}

@@

ggplot(responses_item, aes(reorder(word_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
  facet_wrap(~group, scales = "free_y") +
  geom_hline(yintercept = 0.5, colour = "black") +
  geom_errorbar(width = 0) +
  geom_point(size = 0.5) +
  labs(y = "Proportion of correct responses", x = "Word", colour = "Group") +
  coord_flip() +
  scale_color_manual(values = c("#1E88E5", "#ff2976", "#FFC20A")) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  scale_x_discrete(labels = function(x) str_remove(x, "___ENG-CAT|___ENG-SPA|___SPA-CAT")) +
  theme(
    legend.position = "none",
    axis.text.y = element_text(size = 5, colour = "black", hjust = 1),
    axis.ticks = element_blank(),
    axis.text.x = element_text(size = 7, colour = "black"),
  )

>>>>

bibliography:
  - r-references.bib
  - references.bib
floatsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: no
draft: no
csl: "apa7.csl"
documentclass: apa7
classoption: man
output: papaja::apa6_docx
---

@@

options(
  knitr.kable.NA = '-', 
  knitr.duplicate.label = "allow",
  gt.html_tag_check = FALSE
  )

@@

$$
\begin{align*}

&\textbf{Likelihood}  \\
y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\

&\textbf{Parameters}  \\

logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
& \beta_{1[p]} ~ Frequency_{i} ~ + \\
& \beta_{2[p]} ~ PTHN_i ~ + \\
& \beta_{3[p]} ~ Similarity_i ~ + \\
& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\

\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
&& \text{[participant-level coefficients]} \\ \\

&\textbf{Prior}  \\

\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 3) && \text{[participant-level coefficients]} \\
\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ \text{half}- Cauchy(0, 3) && \text{[SD for population and participant]} \\
\rho_{p}, ~ \rho_{w} \sim& ~LKJ(5) && \text{[correlation between participant-level coefficients]} \\

\end{align*}
$$

@@

# Appendix 1: MCMC diagnostics

```{r appendix-diagnostics, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

suppressWarnings({
  bayesplot_theme_set(
    new = theme_custom() + 
      theme(
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 8),
        strip.text = element_text(size = 10),
        legend.key = element_rect(colour = NA, fill = "white")
      )
  )
})

color_scheme_set("brightblue")
pars <- c("b_Intercept", "b_frequency_zipf", "b_pthn", "b_lv", "b_pthn:lv")

## Pairs plot

```{r appendix-mcmc, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
posterior_cp <- as.array(model_fits$fit_3)
np_cp <- nuts_params(model_fits$fit_3)

plot_pairs <- mcmc_pairs(
  posterior_cp, pars = pars,
  off_diag_args = list(size = 0.75)
)

plot_pairs
```

## Trace and Rank plots

```{r appendix-trace, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
(mcmc_trace(model_fits$fit_3, pars) + legend_none()) /
  mcmc_rank_overlay(model_fits$fit_3, pars = pars) +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "top"
  )
```

## Gelman-Rubin convergence criterion (R-hat)

```{r appendix-rhat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
rhats <- rhat(model_fits$fit_3)
ratios <- neff_ratio(model_fits$fit_3)

mcmc_rhat_hist(rhats, binwidth = 0.00025, color = "white") + 
  mcmc_neff_hist(ratios, binwidth = 0.1) +
  plot_layout(guides = "keep", ncol = 1) &
  theme(
    legend.position = "top"
  )
```

# Appendix 2: by item accuracy

```{r appendix-accuracy, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Proportion of correct translation by item. Words presented to participants in the English-Spanish or in the English-Catalan and Spanish-Catalan are listed in the Y-axis, ordered from higher to lower average translation accuracy, depicted in the X-axis. Dots and whiskers represent the average accuracy and 95% confidence interval of each word. Accuracy is plotted separately for each group."}

@@

ggplot(responses_item, aes(reorder(word_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
  facet_wrap(~group, scales = "free_y") +
  geom_hline(yintercept = 0.5, colour = "black") +
  geom_errorbar(width = 0) +
  geom_point(size = 0.5) +
  labs(y = "Proportion of correct responses", x = "Word", colour = "Group") +
  coord_flip() +
  scale_color_manual(values = c("#1E88E5", "#ff2976", "#FFC20A")) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  scale_x_discrete(labels = function(x) str_remove(x, "___ENG-CAT|___ENG-SPA|___SPA-CAT")) +
  theme(
    legend.position = "none",
    axis.text.y = element_text(size = 5, colour = "black"),
    axis.ticks = element_blank(),
    axis.text.x = element_text(size = 9, colour = "black"),
    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
  )
